single_elimination

### 2025-02-06 single_elimination_shiny.qmd
update 2/6/2025

@@

```{r}
generate_tournament_structure <- function(n) {

@@

@@

  return(out)
}

@@

```{r}
# Simulate the single-elimination tournament
simulate_tournament <- function(num_teams,distribution,seeding_structure = NULL, ties=T, series=1, third_place = T) {

if (num_teams <= 3) {
  stop("Number of Teams must be greater than 3.")
}
teams <- paste("Team",num_teams:1)
if (log2(num_teams) %% 1 != 0) {
  power_of_2 <- 2^ceiling(log2(num_teams))
  teams <- c(teams, rep(NA, power_of_2 - num_teams))
}
if (distribution == "Normal"){
  strengths <- sapply(num_teams, function(n) { qnorm(1:n/(n+1)) })
  normal_strengths <- data.frame(
  Team = teams,
  Strength = c(strengths, rep(NA, length(teams) - num_teams)),
  Wins = rep(0,length(teams)),
  Ranks = rep(NA,length(teams))
)
normal_strengths$True_Rank <- rank(-normal_strengths$Strength, ties.method = "average")
df <- arrange(normal_strengths, True_Rank)
}
else if (distribution == "Uniform"){
  strengths <- sapply(num_teams, function(n) { qunif(1:n/(n+1) , 0 , sqrt(12)) })
  unif_strength <- data.frame(
  Team = teams,
  Strength = c(strengths, rep(NA, length(teams) - num_teams)),
  Wins = rep(0,length(teams)),
  Ranks = rep(NA,length(teams))
)
unif_strength$True_Rank <- rank(-unif_strength$Strength, ties.method = "average")
df <- arrange(unif_strength, True_Rank)
}
else {
  stop("Distribution must be Normal or Uniform")
}

  # Sets seed to that of common best vs worst structure
  if (is.null(seeding_structure)){
    seeding_structure <- matrix(generate_tournament_structure(num_teams), nrow = 1)
  } else{
    seeding_structure <- matrix(seeding_structure, nrow = 1)
  }
  if(series %% 2 == 0) {
    stop("Series must be an odd number to avoid potential ties.")
  }

@@

  for (i in 1:nrow(seeding_structure)) {
    # Create a new data frame based on the current permutation in 'test'
    permutation <- seeding_structure[i, ]
    permuted_df <- df
    permuted_df$Team <- df$Team[permutation]  # Update team order 
    permuted_df$Strength <- df$Strength[permutation]
    permuted_df$Ranks <- NA  # Initialize ranks column
    permuted_df$Wins <- 0  # Initialize wins column
    permuted_df$True_Rank <- df$True_Rank[permutation]

    teams <- permuted_df$Team
    strengths <- permuted_df$Strength
    round_number <- ceiling(log2(length(teams)))

@@

      strength1 <- strengths[which(permuted_df$Team == team1)]
      strength2 <- strengths[which(permuted_df$Team == team2)]

      wins_team1 <- 0
      wins_team2 <- 0
      for (game in 1:series){
      match_winner <- simulate_match(team1, team2, strength1, strength2)
      #cat(sprintf("Matchup:  %s vs %s | Strengths: %.2f vs %.2f | Winner: %s\n", 
      #      team1, team2, strength1, strength2, match_winner))
      if (match_winner == team1) {
            wins_team1 <- wins_team1 + 1
          } else {
            wins_team2 <- wins_team2 + 1
          }

@@

@@

      # Function to calculate the rank values for a given round
    get_rank_values <- function(n) {
      # Calculate the rank values as the average of the last n/2 teams
      suppressWarnings({
        if (n / 2 + 1 == num_teams) {
          rank_values <- num_teams
        } else {
          rank_values <- (n / 2 + 1):min(num_teams, n)
        }
      })
      return(rank_values)
    }
      if (round_number > 2){
      # Assign ranks based on the current round
        n <- 2^(round_number)
        rank_values <- suppressWarnings(get_rank_values(n))
        #print(match_winner)
        if (match_winner == team1) {
          if (ties == TRUE){
          permuted_df$Ranks[permuted_df$Team == team2] <- suppressWarnings(mean(rank_values))
          }
          else if (ties == FALSE){
          permuted_df$Ranks[permuted_df$Team == team2] <- n
          }
        } else {
          if (ties == TRUE){
          permuted_df$Ranks[permuted_df$Team == team1] <-  suppressWarnings(mean(rank_values))
          }
          else if (ties == FALSE){
          permuted_df$Ranks[permuted_df$Team == team1] <- n
          }
        }
      } else if (round_number == 2) {  # Semifinals
        n <- 2^(round_number)
        rank_values <- get_rank_values(n)
        if (match_winner == team1) {
          if (ties == TRUE){
          permuted_df$Ranks[permuted_df$Team == team2] <-
          mean(rank_values)
          }
          else if (ties == FALSE & third_place == FALSE){
          permuted_df$Ranks[permuted_df$Team == team2] <- n
          }
          losers_semis <- c(losers_semis, team2)
        } else {
          if (ties == TRUE){
          permuted_df$Ranks[permuted_df$Team == team1] <-
          mean(rank_values)
          }
          else if (ties == FALSE & third_place == FALSE){
          permuted_df$Ranks[permuted_df$Team == team1] <- n
          }
          losers_semis <- c(losers_semis, team1)

          }
        }else if (round_number == 1) { # Finals
          permuted_df <- permuted_df %>%
            filter(!is.na(Team))
           n <- 2^(round_number)
        rank_values <- get_rank_values(n)
        if (match_winner == team1) {
          permuted_df$Ranks[permuted_df$Team == team2] <- 2
        permuted_df$Ranks[permuted_df$Team == team1] <- 1
        permuted_df$Wins[permuted_df$Team == team1] <-
        permuted_df$Wins[permuted_df$Team == team1] + 1
        } else {
          permuted_df$Ranks[permuted_df$Team == team1] <- 2
          permuted_df$Ranks[permuted_df$Team == team2] <- 1
          permuted_df$Wins[permuted_df$Team == team2] <-
          permuted_df$Wins[permuted_df$Team == team2] + 1
        }
      }
    }

@@

  # Third place match between the losers of the semifinals
  if (third_place == TRUE){
    team1 <- losers_semis[1]
    team2 <- losers_semis[2]
    strength1 <- permuted_df$Strength[permuted_df$Team == team1]
    strength2 <- permuted_df$Strength[permuted_df$Team == team2]
    wins_team1 <- 0
    wins_team2 <- 0
    for (game in 1:series) {
      match_winner <- simulate_match(team1, team2, strength1, strength2)
      if (match_winner == team1) {
        wins_team1 <- wins_team1 + 1
      } else {
        wins_team2 <- wins_team2 + 1
      }
     # Determine the overall winner if a majority is reached
      if (wins_team1 > series / 2 || wins_team2 > series / 2) {
      break
      }
    }
   # Assign third place and fourth place ranks
      if (wins_team1 > wins_team2) {
        #cat(sprintf("Third Place Matchup: %s vs %s | Winner: %s\n", team1, team2, team1))
        permuted_df$Ranks[permuted_df$Team == team1] <- 3
        permuted_df$Ranks[permuted_df$Team == team2] <- 4
      } else {
        #cat(sprintf("Third Place Matchup: %s vs %s | Winner: %s\n", team1, team2, team2))
        permuted_df$Ranks[permuted_df$Team == team2] <- 3
        permuted_df$Ranks[permuted_df$Team == team1] <- 4
      }
    }

  results[[i]] <- permuted_df # Store the results of this permutation
  }

@@

  # Generalized function to remove ties by assigning unique ranks
remove_ties <- function(df) {
  suppressWarnings({
  df <- df %>%
    group_by(Ranks) %>%
    mutate(
      Ranks = ifelse(
        Ranks == 2,  # Do not change if rank is 2
        Ranks, 
        ifelse(
          log2(Ranks) %% 1 == 0,  # Check if Ranks is a power of 2
          ifelse(
            Ranks / 2 + 1 == num_teams,  
            num_teams,  
            sample((Ranks / 2 + 1):(min(Ranks, num_teams)), n(), replace = FALSE)  # Regular case
          ),
          Ranks  # Keep Ranks as is if it's not a power of 2
        )
      )
    ) %>%
    ungroup()
  })
  return(df)
}

  # Remove ties if ties = FALSE
  if (!ties) {
    final_results <- remove_ties(final_results)
  }

  final_results <- final_results[order(final_results$Ranks),]
  return(final_results)
}

test <- simulate_tournament(num_teams = 10, distribution = "Normal", ties=F, series=3, third_place=T, seeding_structure = c(1,NA,2,NA,3,NA,4,5,6,7,8,NA,9,NA,10,NA))
test

@@

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 4, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 3 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
      actionButton("run_sim", "Run Simulation")
    ),

@@

# Define the server logic
server <- function(input, output) {
  custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_tournament_structure(input$num_teams))
    }
  })

@@

    simulate_tournament(
      num_teams = input$num_teams,
      distribution = input$distribution,
      ties = input$ties,
      series = input$series,
      third_place = input$third_place,
      seeding_structure = seed_structure
    )
  })

  output$ranked_results <- renderTable({
    results <- tournament_results()
    results[order(-results$Strength), ]  # Order by Strength descending
  })
}

@@

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 3 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
      actionButton("run_sim", "Run Simulation")
    ),

@@

# Shiny server logic
server <- function(input, output) {
  custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_tournament_structure(input$num_teams))
    }
  })

  simulation_results <- eventReactive(input$run_sim, {
    req(input$num_teams > 3, input$series %% 2 == 1)  # Ensure valid inputs
    seed_structure <- custom_seeding()
    # Run simulations
    replicate_results <- replicate(
      input$num_simulations,
      simulate_tournament(
        num_teams = input$num_teams,
        distribution = input$distribution,
        ties = input$ties,
        series = input$series,
        third_place = input$third_place,
        seeding_structure = seed_structure
      ),
      simplify = FALSE
    )

    # Combine all results and summarize total wins
    all_results <- do.call(rbind, replicate_results)
    total_wins_summary <- all_results %>%
      group_by(Team) %>%
      summarise(
        Total_Wins = sum(Wins, na.rm = TRUE),
        .groups = 'drop'
      ) %>%
      arrange(desc(Total_Wins))
    return(total_wins_summary)
  })

@@

    ggplot(results, aes(x = reorder(Team, -Total_Wins), y = Total_Wins, fill = Total_Wins)) +
      geom_text(aes(label = Total_Wins), vjust = -0.5, size = 4, color = "black") +
      geom_bar(stat = "identity") +
      labs(
        title = "Total Wins Across Simulations",
        x = "Team",
        y = "Total Wins"
      ) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_fill_gradient(low = "skyblue", high = "blue")
  })
}

@@

```{r}
library(shiny)

ui <- fluidPage(
  titlePanel("Tournament Simulation"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", 8, min = 4, max = 32, step = 1),
      selectInput("distribution", "Distribution Type:", choices = c("Normal", "Uniform")),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 3 7"),
      numericInput("replicates", "Number of Replicates:", 100, min = 1, max = 1000, step = 1),
      numericInput("series", "Number of Games in a Series (odd number):", 5, min = 3, max = 11, step = 2),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
      actionButton("run_sim", "Run Tournament Simulation")
    ),

@@

server <- function(input, output, session) {
 custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_tournament_structure(input$num_teams))
    }
  })

@@

    # Run tournament simulations for the specified number of replicates
    simulation_results <- list()
    for (replicate in 1:input$replicates) {
      simulation_results[[replicate]] <- simulate_tournament(
        num_teams = input$num_teams,
        distribution = input$distribution,
        ties = input$ties,
        series = input$series,
        third_place = input$third_place,
        seeding_structure = seed_structure
      )

@@

      # Compute average ranks across all replicates
      average_ranks <- all_results %>%
        group_by(Team, True_Rank) %>%
        summarise(Average_Rank = mean(Ranks, na.rm = TRUE), .groups = 'drop') %>%
        arrange(Team)

      # Calculate Kendall Tau correlation for the current replicate
      kendall_correlations[replicate] <- cor(average_ranks$Average_Rank, average_ranks$True_Rank, method = "kendall", use = "complete.obs")

      # Calculate Spearman correlation for the current replicate
      spearman_correlations[replicate] <- cor(average_ranks$Average_Rank, average_ranks$True_Rank, method = "spearman", use = "complete.obs")
    }

@@

```{r}
ui <- fluidPage(
  titlePanel("Tournament Simulation with Total Wins Visualization"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 3 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
      actionButton("run_sim", "Run Simulation")
    ),

>>>>

### Things To Do

-   Add a re-seeded option

-   Custom strengths (likely as an input to the function)

@@

```{r}
generate_elimination_tournament_structure <- function(n) {

@@

  left_half <- teams[1:(length(teams)/2)]
  right_half <- teams[(length(teams)/2+1):length(teams)]
  right_half <- rev(right_half)
  teams <- c(left_half,right_half)

  # Ensure that the lower number team is on the left side of each matchup
  teams <- unlist(lapply(seq(1, length(teams), by = 2), function(i) {
    if (teams[i] > teams[i + 1]) {
      c(teams[i + 1], teams[i])  # Swap if needed
    } else {
      c(teams[i], teams[i + 1])
    }
  })
  )

@@

  return(out)
}
generate_elimination_tournament_structure(8)
generate_elimination_tournament_structure(16)

@@

```{r}
# Simulate the single-elimination tournament
simulate_single_elimination_tournament <- function(num_teams,distribution,seeding_structure = NULL, ties=T, series=1, third_place = T) {

  if (num_teams <= 3) {
    stop("Number of Teams must be greater than 3.")
  }
  teams <- paste(num_teams:1)
  if (log2(num_teams) %% 1 != 0) {
    power_of_2 <- 2^ceiling(log2(num_teams))
    teams <- c(teams, rep(NA, power_of_2 - num_teams))
  }
  if (distribution == "Normal"){
    strengths <- sapply(num_teams, function(n) { qnorm(1:n/(n+1)) })
    normal_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA,length(teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams))
    )
    df <- arrange(normal_strengths, true_rank)
  }
  else if (distribution == "Uniform"){
    strengths <- sapply(num_teams, function(n) { qunif(1:n/(n+1) , 0 , sqrt(12)) })
    unif_strength <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams)),
      rank_hat = rep(NA,length(teams))
    )
    df <- arrange(unif_strength, true_rank)
  }
  else if (distribution == "Manual") {
    cat("Please enter the strengths for each team in descending rank order:\n")
    strengths <- numeric(num_teams)
    for (i in num_teams:1) {
      repeat {
        input <- as.numeric(readline(paste("Strength for Team", num_teams - i + 1, ": ")))
        if (!is.na(input)) {
          strengths[i] <- input
          break
        } else {
          cat("Invalid input. Please enter a numeric value.\n")
        }
      }
    }
    manual_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA, length(teams)),
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams))
    )
    df <- arrange(manual_strengths, true_rank)
  }
  else {
    stop("Distribution not found: Enter Manual to input your own strengths")
  }

  # Sets seed to that of common best vs worst structure
  if (is.null(seeding_structure)){
    seeding_structure <- matrix(generate_elimination_tournament_structure(num_teams), nrow = 1)
  } else{
    seeding_structure <- matrix(seeding_structure, nrow = 1)
  }
  if(series %% 2 == 0) {
    stop("Series must be an odd number to avoid potential ties.")
  }

@@

  for (i in 1:nrow(seeding_structure)) {
    # Create a new data frame based on the current permutation in 'test'
    permutation <- seeding_structure[i, ]
    permuted_df <- df
    permuted_df$true_rank <- df$true_rank[permutation]  # Update team order 
    teams <- permuted_df$true_rank
    permuted_df$true_strength <- df$true_strength[permutation]
    permuted_df$rank_hat <- NA  # Initialize ranks column
    permuted_df$game_wins <- 0  # Initialize wins column
    permuted_df$game_losses <- 0

    strengths <- permuted_df$true_strength
    round_number <- ceiling(log2(length(teams)))

@@

        strength1 <- strengths[which(permuted_df$true_rank == team1)]
        strength2 <- strengths[which(permuted_df$true_rank == team2)]

        wins_team1 <- 0
        wins_team2 <- 0
        losses_team1 <- 0
        losses_team2 <- 0
        for (game in 1:series){
          match_winner <- simulate_match(team1, team2, strength1, strength2)
          #cat(sprintf("Matchup:  %s vs %s | Strengths: %.2f vs %.2f | Winner: %s\n", team1, team2, strength1, strength2, match_winner))
          if(!is.na(team1)) {
          if (match_winner == team1) {
            wins_team1 <- wins_team1 + 1
            losses_team2 <- losses_team2 + 1
          } else {
            wins_team2 <- wins_team2 + 1
            losses_team1 <- losses_team1 + 1
          }
          }
          else {wins_team2 <- series / 2 + 1}

@@

        # adds wins and losses of the series to the teams unless they have a bye
        if (!any(is.na(c(team1, team2)))) {
          permuted_df <- permuted_df %>%
            mutate(game_wins = ifelse(true_rank == team1, game_wins + wins_team1, game_wins)) %>%
            mutate(game_wins = ifelse(true_rank == team2, game_wins + wins_team2, game_wins)) %>%
            mutate(game_losses = ifelse(true_rank == team1, game_losses + losses_team1, game_losses)) %>%
            mutate(game_losses = ifelse(true_rank == team2, game_losses + losses_team2, game_losses)) 
           }

@@

        if (round_number > 2){
          # Assign ranks based on the current round
          n <- 2^(round_number)
          #print(series_winner)
            if (!is.na(team1) & series_winner == team1) {
                permuted_df$rank_hat[permuted_df$true_rank == team2] <- n
            } else {
                permuted_df$rank_hat[permuted_df$true_rank == team1] <- n
            }
        } else if (round_number == 2) {  # Semifinals
          n <- 2^(round_number)
          if (match_winner == team1) {
              permuted_df$rank_hat[permuted_df$true_rank == team2] <- n
            losers_semis <- c(losers_semis, team2)
          } else {
              permuted_df$rank_hat[permuted_df$true_rank == team1] <- n
            losers_semis <- c(losers_semis, team1)
          }
        }else if (round_number == 1) { # Finals
          permuted_df <- permuted_df %>%
            filter(!is.na(true_rank))
          n <- 2^(round_number)
          if (series_winner == team1) {
            permuted_df$rank_hat[permuted_df$true_rank == team2] <- 2
            permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
          } else {
            permuted_df$rank_hat[permuted_df$true_rank == team1] <- 2
            permuted_df$rank_hat[permuted_df$true_rank == team2] <- 1
          }
        }
      }

@@

    # Third place match between the losers of the semifinals
    if (third_place == TRUE){
      team1 <- losers_semis[1]
      team2 <- losers_semis[2]
      strength1 <- permuted_df$true_strength[permuted_df$true_rank == team1]
      strength2 <- permuted_df$true_strength[permuted_df$true_rank == team2]
      wins_team1 <- 0
      wins_team2 <- 0
      for (game in 1:series) {
        match_winner <- simulate_match(team1, team2, strength1, strength2)
        if (match_winner == team1) {
          wins_team1 <- wins_team1 + 1
        } else {
          wins_team2 <- wins_team2 + 1
        }
        # Determine the overall winner if a majority is reached
        if (wins_team1 > series / 2 || wins_team2 > series / 2) {
          break
        }
      }
      # Assign third place and fourth place ranks
      if (wins_team1 > wins_team2) {
        #cat(sprintf("Third Place Matchup: %s vs %s | Winner: %s\n", team1, team2, team1))
        permuted_df$rank_hat[permuted_df$true_rank == team1] <- 3
        permuted_df$rank_hat[permuted_df$true_rank == team2] <- 4
      } else {
        #cat(sprintf("Third Place Matchup: %s vs %s | Winner: %s\n", team1, team2, team2))
        permuted_df$rank_hat[permuted_df$true_rank == team2] <- 3
        permuted_df$rank_hat[permuted_df$true_rank == team1] <- 4
      }
    }

    permuted_df$seed <- rep(paste(seeding_structure[i, ], collapse = ", "), nrow(permuted_df))
    results[[i]] <- permuted_df # Store the results of this permutation
  }

@@

  # Remove ties if ties = FALSE
  if (!ties) {
    final_results <- final_results %>%
      filter(!is.na(true_rank)) %>%
      arrange(true_rank) %>%
      mutate(rank_hat = rank(rank_hat, ties.method = "first"))
  } else{
    final_results <- final_results %>%
      filter(!is.na(true_rank)) %>%
      arrange(rank_hat) %>%
      mutate(rank_hat = rank(rank_hat, ties.method = "average"))
  }

  final_results <- final_results %>%
    mutate(distribution = distribution) %>%
    mutate(true_rank = as.numeric(true_rank)) %>%
    arrange(true_rank)
  return(final_results)
}

test <- simulate_single_elimination_tournament(num_teams = 8, distribution = "Normal", ties=F, series=1, third_place=T)
test

#manual_test <- simulate_single_elimination_tournament(num_teams = 4, distribution = "Manual", ties=T, series=1, third_place=T)
#manual_test

# Gets all results of a normal distribution num_teams between 4 and 64 single-game elimination
all_normal_results <- list()
for (teams in c(4,8,16,32,64)) {
  all_normal_results[[as.character(teams)]] <- list()
  for (replicate in 1:10000) {
    # Store each simulation result in the list with unique indexing
    all_normal_results[[as.character(teams)]][[replicate]] <- simulate_single_elimination_tournament(
      num_teams = teams,
      distribution = "Normal",
      ties = FALSE,
      series = 1,
      third_place = TRUE
    )
  }
}

all_results_norm <- do.call(rbind, lapply(names(all_normal_results), function(teams) {
  team_results <- do.call(rbind, all_normal_results[[teams]])
  team_results$teams <- as.numeric(teams)
  return(team_results)
}))

all_results_norm

# Gets all results of a uniform distribution num_teams between 4 and 64 single-game elimination
all_unif_results <- list()
for (teams in c(4,8,16,32,64)) {
  all_unif_results[[as.character(teams)]] <- list()
  for (replicate in 1:100) {
    # Store each simulation result in the list with unique indexing
    all_unif_results[[as.character(teams)]][[replicate]] <- simulate_single_elimination_tournament(
      num_teams = teams,
      distribution = "Uniform",
      ties = FALSE,
      series = 1,
      third_place = TRUE
    )
  }
}

all_results_unif <- do.call(rbind, lapply(names(all_unif_results), function(teams) {
  team_results <- do.call(rbind, all_unif_results[[teams]])
  team_results$teams <- as.numeric(teams)
  return(team_results)
}))

all_results_unif

@@

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 4, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      uiOutput("manual_strengths_ui"), # Dynamic UI for Manual strengths
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
      actionButton("run_sim", "Run Simulation")
    ),

@@

# Define the server logic
server <- function(input, output) {
  custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_elimination_tournament_structure(input$num_teams))
    }
  })

@@

    simulate_single_elimination_tournament(
      num_teams = input$num_teams,
      distribution = input$distribution,
      ties = input$ties,
      series = input$series,
      third_place = input$third_place,
      seeding_structure = seed_structure
    )
  })

  output$ranked_results <- renderTable({
    results <- tournament_results()
    results[order(results$true_rank), ]  # Order by Strength descending
  })
}

@@

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
      actionButton("run_sim", "Run Simulation")
    ),

@@

# Shiny server logic
server <- function(input, output) {
  custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_elimination_tournament_structure(input$num_teams))
    }
  })

  simulation_results <- eventReactive(input$run_sim, {
    req(input$num_teams > 3, input$series %% 2 == 1)  # Ensure valid inputs
    seed_structure <- custom_seeding()
    # Run simulations
    replicate_results <- replicate(
      input$num_simulations,
      simulate_single_elimination_tournament(
        num_teams = input$num_teams,
        distribution = input$distribution,
        ties = input$ties,
        series = input$series,
        third_place = input$third_place,
        seeding_structure = seed_structure
      ),
      simplify = FALSE
    )

    # Combine all results and summarize total wins
    all_results <- do.call(rbind, replicate_results)
    total_wins_summary <- all_results %>%
      group_by(true_rank) %>%
      summarise(
        total_wins = sum(rank_hat==1, na.rm = TRUE),
        .groups = 'drop'
      ) %>%
      arrange(desc(total_wins))
    return(total_wins_summary)
  })

@@

    ggplot(results, aes(x = reorder(true_rank, -total_wins), y = total_wins, fill = total_wins)) +
      geom_text(aes(label = total_wins), vjust = -0.5, size = 4, color = "black") +
      geom_bar(stat = "identity") +
      labs(
        title = "Total Wins Across Simulations",
        x = "True Rank",
        y = "Total Wins"
      ) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      scale_fill_gradient(low = "skyblue", high = "blue")
  })
}

@@

```{r}
ui <- fluidPage(
  titlePanel("Tournament Simulation"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", 8, min = 4, step = 1),
      selectInput("distribution", "Distribution Type:", choices = c("Normal", "Uniform")),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
      numericInput("replicates", "Number of Replicates:", 100, min = 1, max = 1000, step = 1),
      numericInput("series", "Number of Games in a Series (odd number):", 1, min = 1, max = 11, step = 2),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
      actionButton("run_sim", "Run Tournament Simulation")
    ),

@@

server <- function(input, output, session) {
 custom_seeding <- reactive({
    if (input$seeding_input != "") {
      # Convert the comma-separated string into a numeric vector
      seeding_vector <- unlist(strsplit(input$seeding_input, " "))
      seeding_vector[seeding_vector == "NA"] <- NA
      seeding_vector <- as.numeric(seeding_vector)
      num_teams_next_power_of_two <- 2^ceiling(log2(input$num_teams))
      # Validate that it matches the number of teams
      if (length(seeding_vector) == num_teams_next_power_of_two) {
        return(seeding_vector)
      } else {
        showNotification(paste("Seeding structure length must match", num_teams_next_power_of_two), type = "error")
        return(NULL)
      }
    } else {
      return(generate_elimination_tournament_structure(input$num_teams))
    }
  })

@@

    # Run tournament simulations for the specified number of replicates
    simulation_results <- list()
    for (replicate in 1:input$replicates) {
      simulation_results[[replicate]] <- simulate_single_elimination_tournament(
        num_teams = input$num_teams,
        distribution = input$distribution,
        ties = input$ties,
        series = input$series,
        third_place = input$third_place,
        seeding_structure = seed_structure
      )

@@

      # Compute average ranks across all replicates
      average_ranks <- all_results %>%
        group_by(true_rank) %>%
        summarise(avg_rank = mean(rank_hat, na.rm = TRUE), .groups = 'drop') %>%
        arrange(true_rank)

      # Calculate Kendall Tau correlation for the current replicate
      kendall_correlations[replicate] <- cor(all_results$rank_hat, all_results$true_rank, method = "kendall", use = "complete.obs")

      # Calculate Spearman correlation for the current replicate
      spearman_correlations[replicate] <- cor(all_results$rank_hat, all_results$true_rank, method = "spearman", use = "complete.obs")
    }

@@

```{r}
ui <- fluidPage(
  titlePanel("Probability of Top 3 Teams in the Correct Order"),

  sidebarLayout(
    sidebarPanel(
      numericInput("num_teams", "Number of Teams:", value = 8, min = 2, step = 1),
      textInput("seeding_input", "Custom Seeding Structure (space-separated, NA for byes):", placeholder = "e.g. 1 8 4 5 3 6 2 7"),
      selectInput("distribution", "Strength Distribution:", choices = c("Normal", "Uniform")),
      numericInput("series", "Best of Series (Odd Number):", value = 1, min = 1, step = 2),
      numericInput("num_simulations", "Number of Simulations:", value = 100, min = 10, step = 10),
      checkboxInput("ties", "Allow Ties?", value = TRUE),
      checkboxInput("third_place", "Include Third Place Match?", value = TRUE),
      actionButton("run_sim", "Run Simulation")
    ),
